A community-maintained standard library of population genetic simulations

### 2019-12-18 popSimManu.tex
Various edits for clarity and grammar.

\newcommand{\stdpopsim}{\texttt{stdpopsim}\xspace}
\newcommand{\dadi}{$\partial a \partial i$\xspace}
\newcommand{\MSMC}{\texttt{MSMC}\xspace}
\newcommand{\smcpp}{\texttt{smc++}\xspace}
\newcommand{\stairwayplot}{\texttt{stairwayplot}\xspace}
\newcommand{\fastsimcoal}{\texttt{fastsimcoal}\xspace}

@@

\section*{Introduction}
While population genetics has always used statistical methods to make inferences from data,
the degree of sophistication of the questions, models, data, and computational approaches
used have all increased over the past decade. Currently there exist myriad computational methods
that can infer the histories of populations, the distribution of fitness effects,
recombination rates, the extent of positive selection in genome sequence data
\citep{li2011inference,schiffels2014inferring,terhorst2017robust,eyre2009estimating,kim2017inference,
chan2012genome,lin2013fast,Adrion662247,alachiotis2012omegaplus,degiorgio2016sweepfinder2,
kern2018diplos,sugden2018localization}.
While en masse these methods have increased our understanding of the
impacts of genetic and evolutionary processes, very little has been done to systematically
benchmark the quality of inferences gleaned from computational population genetics.
As large databases of population genetic variation begin to be used to inform public health procedures,
the accuracy and quality of these inferences is becoming even more important.
This is in large part because ``ground-truth'' in population genetics comes from simulation,
not from empirical observations, and as a field there has been no agreement as to what should
constitute community standards or best practices. The general modus operandi to date has been to
introduce methods which are validated by a set of bespoke simulations produced by individual
groups. Beyond a lack of efficiency from duplicated effort, such a situation leads
to decreases in reproducibility and transparency across the entire field.
Here we describe \stdpopsim, a community-driven catalog
of standardized population genetics models spanning a variety of common study organisms.
This resource will empower the field to rigorously evaluate population genetic software
and select the best tool for a given task, thereby helping to usher in the next era of population genetic data analysis.
%DRS: still not in love with this last sentence here, but tried to make it a little less sappy by adding some explanation.
%ADK: made some edits here

@@

One way to benchmark different methods for statistical inference in population
genetics is to apply the methods to simulated datasets where the true model
parameters are known. Methods that perform well will accurately infer these
parameters and will also provide reasonable estimates of uncertainty of the
estimated parameters. Indeed, simulations studies are a standard ingredient in
most papers proposing new methods in population genetics. However, these
studies often use simulations of the particular models for which the method was designed,
which showcases the novel method rather than rigorously comparing performance to competing methods.
Of course, testing a new method in the situation for which it was designed is an important step,
but more rigorous testing is necessary for tools that are to be widely used in the field.
% Moreover researchers are generally forced to reimplement previously
% inferred scenarios independently, which makes good benchmarking
% a major additional time investment, and can lead to errors.
% Lastly, the degree to which different methods are compared can be quite variable.
Many more researchers in the broader field
use simulations to perform power and other exploratory analyses.
The many steps of implementing a moderately complicated population genetics simulation,
including translating published demographic models into input for a simulator
and obtaining genetic maps,
present a major obstacle, especially to new researchers.

@@

For these reasons, we have generated a standardized, community-driven resource
for simulating published demographic models from a number of popular study systems.
This resource, which we call \texttt{stdpopsim}, makes running
realistic simulations for population genetic analysis a simple matter of
choosing pre-implemented models from a community maintained catalog.
The \stdpopsim catalog currently contains four organisms: humans, 
\emph{D.~melanogaster}, \emph{A.~thaliana}, and \emph{E.~coli}. For each
organism the catalog contains details on the physical organization (e.g., chromosome structure)
of their genomes, one or more genetic maps, default population-level parameters (mutation rate,
generation time) and one or more published demographic histories. Through
either a command line interface or a simple Python API, the user can specify which
organism, genetic map, chromosome, and demographic history they are interested in simulating, and are returned
simulation output from their chosen model.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.5\linewidth]{display_items/Figure1.pdf}
\caption{\textbf{Structure of \stdpopsim}. \textbf{A.} The
hierarchical organization of the PopSim catalog contains all information
within individual species. Information within the species level is shown here
for \emph{Homo sapiens} only. We see that \emph{H. sapiens} is associated
with a default generation time and mutation rate, that it contains within
it a representation of the physical genome, genetic maps, and as shown a
subset of the demographic models currently implemented. \textbf{B.} Caching
and automated download of genetic maps. \textbf{C.} Example code to specify
and simulate models using \stdpopsim via the API (top) or
command-line interface (bottom). \textbf{D.} Simulation output from
\stdpopsim is generated as tree sequence files that can
be manipulated and analysed with \texttt{tskit}.}
\label{fig:cartoon}
\end{center}
\end{figure}

\section*{Results}
\paragraph{The \stdpopsim library}
The central contribution of the PopSim consortium currently is \stdpopsim, a community
maintained library of empirical genome data and population genetics simulation
models, providing easy access to realistic simulations. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig. \ref{fig:cartoon}A), initially consisting of humans, \emph{Drosophila},
\emph{Arabidopsis}, \emph{E. coli}, and \emph{Pongo}. A species definition consists
of two key elements.  Firstly, the library defines
some basic information about its genome, including information about chromosome
lengths, average mutation rates and so on. We also provide access to detailed
empirical information such as recombination maps, which model empirically
observed heterogeneity along chromosomes. As such maps are often large and hard
to find, we provide a mechanism to automatically download (and cache locally) a
set of previously published maps (Fig. \ref{fig:cartoon}B). For example, in humans we support the
HapMapII~\citep{international2007second} and
deCODE~\citep{kong2010fine} genetic maps. The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which are the basis for accurate simulations
of specific historical scenarios that have influenced present-day patterns of
genetic polymorphism.

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run accurate, standardized simulations
across a range of organisms. \stdpopsim has a Python API and a user-friendly
command line interface (Fig. \ref{fig:cartoon}C), allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the “tree sequence”
format \citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}, which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed very efficiently using the tskit library
\citep{kelleher2016efficient,kelleher2018efficient}. Currently,
\stdpopsim uses the  \texttt{msprime} coalescent simulator \citep{kelleher2016efficient}
as the default simulation engine. We plan to include other simulation
engines to support scenarios that cannot be modelled under the coalescent framework,
and have already implemented \texttt{SLiM} \citep{haller2019slim} as
an alternative back-end.

@@

At the time of writing, more models of human history have been implemented than for
any other organism. These models include: a simplified version of the \cite{tennessen2012evolution}
model with only the African population modeled (expansion from the ancestral
population and recent growth), the three population model of \cite{gutenkunst2009inferring}
which modeled the out-of-Africa bottleneck as well as the subsequent divergence of
the European and Asian populations, the \cite{tennessen2012evolution} two population variant of the
Gutenkunst model which did not include Asian populations, but more explicitly modeled
recent rapid human population growth, the \cite{browning2018ancestry} admixture model
for American populations that models ancestral African, European, and Asian population
components, and the three-population out-of-Africa model from \cite{ragsdale2019models} that includes archaic admixture
. Together these models
contain features believed to have widespread impacts in real data (e.g. bottlenecks, population growth,
admixture) which are therefore highly pertinent in the context of method development.

For non-human genomes we have have implemented two demographic histories for
\emph{Drosophila melanogaster} and one each from \emph{A. thaliana} and \emph{E. coli}.
For \emph{D. melanogaster} we have implemented the three-epoch model estimated by \cite{sheehan2016deep} from
an African sample, as well as the out-of-Africa divergence
and associated bottleneck model of \cite{li2006inferring} which jointly models African
and European populations. For \emph{Arabidopsis thaliana}, we implemented the
model from in \cite{durvasula2017african} inferred using \MSMC. This model includes
a continuous change in population size over time, rather than pre-specified epochs of different
population sizes.

\subsection*{Use case: comparing methods of demographic inference}
As an example of the utility of \stdpopsim we demonstrate how popular
methods can be simply and fairly compared in the setting of demographic
inference using our resource. Although we present comparison of results from several
competing methods, our aim at this stage is not to provide an exhaustive
evaluation or ranking of these methods. Our hope is instead that future work built upon this resource
will enable more detailed exploration of the strengths and weaknesses of the numerous
inference methods that are available to the population genetics community.

We start by comparing popular methods for estimating
population size histories ($N(t)$) of single populations and subsequently
show simple examples of multi-population inference.
To evaluate and compare the performance of inference methods we developed
\texttt{snakemake} \citep{koster2012snakemake} workflows that allow reproducibility
in comparing inference methods which are available from \url{https://github.com/popgensims/analysis}
. The workflows allow efficient computing in multicore environments or on
cluster environments.

For single-population population size histories, we compared
\MSMC~\citep{schiffels2014inferring}, \smcpp~\citep{terhorst2017robust}, and
\stairwayplot~\citep{liu2015exploring}
 on simulated genomes sampled from a single population,
in a number of the \stdpopsim models described above. The pipeline was generalized to
run $R$ replicates with $C$ chromosomes for $n$ samples, for a total of $R \times C$
simulations for each demographic model. After simulation, the pipeline prepares
input files for each of the respective inference methods by grouping all
chromosomes, for each sample, into a single file. For each of the $R$ simulation replicates, this step results in an
input file for each of the
respective inference methods and derived from the same simulated tree sequences.
Each of the inference programs are then run in parallel, followed by plotting of
$N(t)$ estimates from each program.

@@

% JK here --- I think it would really help if we described what the models
% were, briefly, rather than saying who inferred them.
In Figure \ref{fig:n_t_ragsdale} we present results from our $N(t)$ analysis pipeline
run on whole genome simulations under a sophisticated demographic model
and empirical genetic map. In each column of this figure
we show inferred $N(t)$ for the three extant populations in the model.
In each row we show comparisons among the methods (note that we compare two differing
sample sizes for \MSMC). The solid black lines represent the true effective population
size as inferred from the inverse coalescent rate defined by the model.
Blue lines represent estimates from each of three replicate whole genome simulations.
While there is variation in accuracy among methods, populations, and individual replicates,
generally spoken the methods are accurate for this model of human history.

\stdpopsim allows us to readily compare these estimates to those drawn from a different
model of human history. In Figure \ref{fig:n_t_gutenkunst} we show estimates of
$N(t)$ from simulations using the same physical and genetic maps, but a different demographic
history -- that inferred by \cite{gutenkunst2009inferring}. Again we see that each
of the methods is capturing relevant parts of the population history, although the
discrepancy between inferred and simulated values is changes over time. In comparing inferences between the
models it is interesting to note that $N(t)$ estimates for the CHB and CEU
simulated populations are generally better across methods than estimates from the YRI
simulated population.

We can even do away with the demographic model entirely, and simulate under a human
genome architecture and genetic map to see how well methods might to at recovering
the population history of a constant sized population. We show results of such an
experiment in Figure \ref{fig:n_t_HomSap_constant}. Again we see that all methods
are recovering a modern population size within a factor of two of the truth, however
SMC-based methods, perhaps due to their regularization, tend to learn sinusoidal
patterns of population size even though no change is present.

@@

As most population genetic methods development has been focused on human-scale
data, it is of consequence to ask how such methods might perform in non-human
genomes. Figure \ref{fig:n_t_sheehan} shows parameter estimates from a demographic
model estimated from an African sample of \emph{Drosophila melanogaster} \citep{sheehan2016deep}.
As with humans, we use \stdpopsim to simulate replicate \emph{Drosophila} genomes under
an empirically derived genetic map, and then try to infer back the parameters of model.
Accuracy is mixed among methods in this setting, and generally worse than what we
observe under simulations of the human genome.

\begin{figure}
\begin{center}
\includegraphics[width=0.7\linewidth]{display_items/homo_sapiens_two_popn_comp.png}
\caption{\textbf{A comparison two-population parameter estimates in humans}. Here we show estimates of $N(t)$ inferred using \dadi, \texttt{fastsimcoal2}, or \smcpp.
A) Data were generated by simulating
replicate human genomes under the \cite{gutenkunst2009inferring} model and using the genetic map
inferred in \cite{international2007second}. B) For \dadi and \texttt{fastsimcoal2} we show inferred
parameters of the depicted IM model, which include population sizes, migration rates, and a split
time between CEU and YRI samples. C) Population size estimates for each population (rows)
from \dadi, \texttt{fastsimcoal2}, and \smcpp (columns). In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model. The population split date, $T_{div}$, is shown at
the bottom, with a common X-axis to the background figure.}
\label{fig:IM_popn_human}
\end{center}
\end{figure}

\paragraph*{Multi Population demographic models}
As \texttt{stdpopsim} implements multi-population demographic models, we also
explored parameter estimation of population divergence parameters. In particular
we performed simulations under multi-population models for humans and \emph{Drosophila}
and inferred simulated parameters using \dadi, \fastsimcoal, and \smcpp.
For simplicity, we conducted inference in \dadi and \fastsimcoal using an IM model
with constant population sizes and bi-directional migration. For human
models with more than two-populations, e.g. \cite{gutenkunst2009inferring},
this means that we are inferring parameters from a model which itself does
not match the model from which the data were generated (Figure
\ref{fig:IM_popn_human}A and B), however this is common practice in the field
when applying IM-type models.

In Figure \ref{fig:IM_popn_human}C we show estimates of population sizes and divergence
time from each of the methods for samples simulated for the African and European populations
under the \cite{gutenkunst2009inferring} model. Each of the methods compared has stronger
and weaker points to their accuracy across parameters. For instance the simple, IM-style
analyses we have performed do not capture recent population size change in CEU, however
they do a good job of estimating the simpler simulated YRI population size history from this model.

Again we can compare between species and look at the performance of these methods in
the context of a two-population model from \emph{Drosophila melanogaster}. Figure
\ref{fig:two_popn_fly} shows parameter estimates from simulations drawn from
the model from \cite{li2006inferring} which includes
an ancestral population in Africa, then a population expansion with a population
split and bottleneck into a European population. These simulations are generated
using the recombination map of \cite{comeron2012many}. Here methods differ dramatically
in the estimates, with \smcpp yielding quite noisy estimates of population size
towards the more ancient past and all methods failing to capture the very brief
population bottleneck in Europe, although the IM-model of course only allows for
two epochs in the population size history.

@@

The \stdpopsim library is the first piece of this effort, in that it allows for rigorous
standardization of complex population genetic simulations. Population genetics, as a field,
has yet to coalesce around a set of gold-standards for the crucial task of method
evaluation, which in our field hinges on simulation. In contrast, other fields such as
structural biology (\citep{moult1995large}) and machine learning (\citep{russakovsky2015imagenet}) have a long track record
of standardized method testing. We hope that our efforts represent the beginning of what
will prove to be an equally longstanding and invaluable tradition in population genetics.

@@

We have built \stdpopsim, and the PopSim Consortium, to be an open, community-developed
project, meaning that it has excellent potential for future growth.
Thus far we have implemented genome representations and genetic maps for the three most
common study systems in computational genetics: humans, \emph{Drosophila}, and \emph{Arabidopsis}.
The standard operating procedures that we have outlined in our detailed documentation
means that new researchers, who are interested in adding their favorite organism/genome
are free to with very few barriers to entry (see Methods). Further new demographic models can be added
to each of the organisms in the catalog through similarly well documented steps.
A long-term goal then is that through a mechanism such as \stdpopsim, the community
can determine (in an open way) what is the correct implementation for a given model,
and in turn, what are the models that the community is most interested in.

@@

In the use case comparisons we have performed,
we found that the methods tested \emph{do} generally perform well on human genomes,
up to, for instance, the usual factor-of-two differences in inferred population sizes.
This will not be surprising to people familiar with the landscape of
demographic inference methods in population genetics,
but the nature and scale of uncertainty may be not clear to people new to the field.
The ability for \stdpopsim to generate data with and without significant features, such
as the genetic map or population size change (e.g., Figure \ref{fig:n_t_HomSap_constant}), allows
us to ask directly about the various failure modes of popular methods.
Moreover the comparison of methods across the various genome organizations, genetic maps,
and demographic histories of different organisms provides valuable information
about how methods might perform on non-human systems.
In our initial comparison we find that inference of population histories from \emph{Drosophila}
or \emph{Arabidopsis} simulated genomes is a fair bit more difficult from SMC based methods
than in the human genome. Likely this is at least partly a consequence of the reduced physical and genetic
map lengths of these genomes, although we will leave further investigation of this to the future.
Finally, comparison of results across methods or simulation runs
gives a strong estimate of inference uncertainty, analogously to parametric bootstrapping --
although different methods have different strengths and weaknesses,
strong conclusions should probably not rest entirely on features inferred by only one method.

The broad brush comparisons we have done, while far from exhaustive, are informative,
and highlight the utility of easy comparison of methods on the same footing.
The tests here were under best-case scenarios --
simulations of discrete, randomly mating populations that fit the assumptions of the methods nearly exactly --
and yet different methods, as well as applications of the same method
to different simulations of the same model,
produced somewhat different results. Future efforts will extend the resource described here to
encompass additional population genetic tasks, such as detecting natural selection, inferring
recombination rates, etc, in hopes of producing a general-purpose resource for fair and
standardized evaluation of the ever-increasing arsenal of computational population genetic
inference tools.

>>>>

\newcommand{\stdpopsim}{\texttt{stdpopsim}\xspace}
\newcommand{\dadi}{$\partial a \partial i$\xspace}
\newcommand{\MSMC}{\texttt{MSMC}\xspace}
\newcommand{\smcpp}{\texttt{smc++}\xspace}
\newcommand{\stairwayplot}{\texttt{stairwayplot}\xspace}
\newcommand{\fastsimcoal}{\texttt{fastsimcoal}\xspace}
\newcommand{\tskit}{\texttt{tskit}\xspace}

@@

\section*{Introduction}
While population genetics has always used statistical methods to make inferences from data,
the degree of sophistication of the questions, models, data, and computational approaches
used have all increased over the past decade. Currently there exist myriad computational methods
that can infer the histories of populations 
\citep{gutenkunst2009inferring,li2011inference,excoffier2013robust,schiffels2014inferring, terhorst2017robust,ragsdale2019models}, 
the distribution of fitness effects \citep{kim2017inference},
recombination rates \citep{chan2012genome,lin2013fast,Adrion662247}, 
the extent of positive selection in genome sequence data
\citep{eyre2009estimating, alachiotis2012omegaplus,degiorgio2016sweepfinder2,kern2018diplos,sugden2018localization}.
While en masse these methods have increased our understanding of the
impacts of genetic and evolutionary processes, very little has been done to systematically
benchmark the quality of inferences gleaned from computational population genetics.
As large databases of population genetic variation begin to be used to inform public health procedures,
the accuracy and quality of these inferences is becoming even more important.
This is in large part because ``ground-truth'' in population genetics comes from simulation,
not from empirical observations, and as a field there has been no agreement as to what should
constitute community standards or best practices. The general modus operandi to date has been to
introduce methods which are validated by a set of bespoke simulations produced by individual
groups. Beyond a lack of efficiency from duplicated effort, such a situation leads
to decreases in reproducibility and transparency across the entire field.
Here we describe \stdpopsim, a community-driven catalog
of standardized population genetics models spanning a variety of common study organisms.
This resource will empower the field to rigorously evaluate population genetic software
and select the best tool for a given task, thereby helping to usher in the next era of population genetic data analysis.
%DRS: still not in love with this last sentence here, but tried to make it a little less sappy by adding some explanation.
%ADK: made some edits here

@@

One way to benchmark different methods for statistical inference in population
genetics is to apply the methods to simulated datasets where the true model
parameters are known. Methods that perform well will accurately infer these
parameters and will also provide reasonable estimates of uncertainty of the
estimated parameters. Indeed, simulation studies are a standard ingredient in
most papers proposing new methods in population genetics. However, these
studies often use simulations of the particular models for which the method was designed,
which showcases the novel method rather than rigorously comparing performance to competing methods.
Of course, testing a new method in the situation for which it was designed is an important step,
but more rigorous testing is necessary for tools that are to be widely used in the field.
% Moreover researchers are generally forced to reimplement previously
% inferred scenarios independently, which makes good benchmarking
% a major additional time investment, and can lead to errors.
% Lastly, the degree to which different methods are compared can be quite variable.
Many more researchers in the broader field
use simulations to perform power and other exploratory analyses.
The many steps of implementing a moderately complicated population genetics simulation,
including translating published demographic models into input for a simulator
and obtaining genetic maps,
present a major obstacle, especially to new researchers.

@@

For these reasons, we have generated a standardized, community-driven resource
for simulating published demographic models from a number of popular study systems.
This resource, which we call \texttt{stdpopsim}, makes running
realistic simulations for population genetic analysis a simple matter of
choosing pre-implemented models from a community maintained catalog.
The \stdpopsim catalog currently contains four organisms: humans, 
\emph{D.~melanogaster}, \emph{A.~thaliana}, and \emph{E.~coli}. For each
organism the catalog contains details on the physical organization (e.g., chromosome structure)
of their genomes, one or more genetic maps, default population-level parameters (mutation rate,
generation time) and one or more published demographic histories. Through
either a command line interface or a simple Python API, the user can specify which
organism, genetic map, chromosome, and demographic history they are interested in simulating, and the
simulation output from their chosen model is returned.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.5\linewidth]{display_items/Figure1.pdf}
\caption{\textbf{Structure of \stdpopsim}. \textbf{A.} The
hierarchical organization of the PopSim catalog contains all information
within individual species. Information within the species level is shown here
for \emph{Homo sapiens} only. We see that \emph{H. sapiens} is associated
with a default generation time and mutation rate,
a representation of the physical genome, and two genetic maps. Also shown is a
subset of the demographic models currently implemented. \textbf{B.} Caching
and automated download of genetic maps. \textbf{C.} Example code to specify
and simulate models using \stdpopsim via the API (top) or
command-line interface (bottom). \textbf{D.} Simulation output
is a tree sequence file that can
be manipulated and analysed with \tskit.}
\label{fig:cartoon}
\end{center}
\end{figure}

\section*{Results}
\paragraph{The \stdpopsim library}
The central contribution of the PopSim consortium currently is \stdpopsim, a
community-maintained library of empirical genome data and population genetics simulation
models, providing easy access to realistic simulations. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig. \ref{fig:cartoon}A), initially consisting of humans, \emph{Drosophila},
\emph{Arabidopsis}, \emph{E. coli}, and \emph{Pongo}. A species definition consists
of two key elements.  Firstly, the library defines
some basic information about the species' genome, including information about chromosome
lengths, average mutation rates and so on. We also provide access to detailed
empirical information such as recombination maps, which model empirically-observed
heterogeneity along chromosomes. As such maps are often large and hard
to find, we provide a mechanism to automatically download (and cache locally) a
set of previously published maps (Fig. \ref{fig:cartoon}B). For example, in humans we support the
HapMapII~\citep{international2007second} and
deCODE~\citep{kong2010fine} genetic maps. The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which are the basis for accurate simulations
of specific historical scenarios that have influenced present-day patterns of
genetic polymorphism.

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run accurate, standardized simulations
across a range of organisms. \stdpopsim has a Python API and a user-friendly
command line interface (Fig. \ref{fig:cartoon}C), allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the “tree sequence”
format \citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}, which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed very efficiently using the \tskit library
\citep{kelleher2016efficient,kelleher2018efficient}. Currently,
\stdpopsim uses the  \texttt{msprime} coalescent simulator \citep{kelleher2016efficient}
as the default simulation engine. We plan to include other simulation
engines to support scenarios that cannot be modelled under the coalescent framework,
and have already implemented \texttt{SLiM} \citep{haller2019slim} as
an alternative back-end.

@@

At the time of writing, more models of human history have been implemented than for
any other organism. These models include: a simplified version of the \cite{tennessen2012evolution}
model with only the African population specified (expansion from the ancestral
population and recent growth), the three population model of \cite{gutenkunst2009inferring}
which specifies the out-of-Africa bottleneck as well as the subsequent divergence of
the European and Asian populations, the \cite{tennessen2012evolution} two population variant of the
Gutenkunst model which does not include Asian populations, but more explicitly models
recent rapid human population growth, the \cite{browning2018ancestry} admixture model
for American populations which specifies ancestral African, European, and Asian population
components, and the three-population out-of-Africa model from \cite{ragsdale2019models}
which includes archaic admixture.
Together these models
contain features believed to have widespread impacts in real data (e.g. bottlenecks, population growth,
admixture), and are therefore highly pertinent in the context of method development.

For non-human genomes we have have implemented two demographic histories for
\emph{Drosophila melanogaster} and one each from \emph{A. thaliana} and \emph{E. coli}.
For \emph{D. melanogaster} we have implemented the three-epoch model estimated by \cite{sheehan2016deep} from
an African sample, as well as the out-of-Africa divergence
and associated bottleneck model of \cite{li2006inferring} which jointly models African
and European populations. For \emph{Arabidopsis thaliana}, we implemented the
model in \cite{durvasula2017african} inferred using \MSMC. This model includes
a continuous change in population size over time, rather than pre-specified epochs of different
population sizes.

\subsection*{Use case: comparing methods of demographic inference}
As an example of the utility of \stdpopsim, we demonstrate how it can be used 
to simply and fairly compare popular demographic inference methods.
Although we present comparison of results from several
competing methods, our aim at this stage is not to provide an exhaustive
evaluation or ranking of these methods. Our hope is instead that future work built upon this resource
will enable more detailed exploration of the strengths and weaknesses of the numerous
inference methods that are available to the population genetics community.

We start by comparing popular methods for estimating
population size histories ($N(t)$) of single populations and subsequently
show simple examples of multi-population inference.
To reproducibly evaluate and compare the performance of inference methods, we developed
\texttt{snakemake} \citep{koster2012snakemake} workflows,
available from \url{https://github.com/popgensims/analysis}.
These workflows allow efficient computing in multicore or cluster environments.

For single-population population size histories, we compared
\MSMC~\citep{schiffels2014inferring}, \smcpp~\citep{terhorst2017robust}, and
\stairwayplot~\citep{liu2015exploring}
 on simulated genomes sampled from a single population,
in a number of the demographic models described above. The pipeline was generalized to
run $R$ replicates with $C$ chromosomes for $n$ samples, for a total of $R \times C$
simulations for each demographic model. After simulation, the pipeline prepares
input files for each of the respective inference methods by grouping all
chromosomes, for each sample, into a single file. For each of the $R$ simulation replicates, this step results in an
input file for each of the
respective inference methods and derived from the same simulated tree sequences.
Each of the inference programs are then run in parallel, followed by plotting of
$N(t)$ estimates from each program.

@@

% JK here --- I think it would really help if we described what the models
% were, briefly, rather than saying who inferred them.
In Figure \ref{fig:n_t_ragsdale} we present results from our $N(t)$ analysis pipeline
run on whole genome simulations under a sophisticated demographic model
and empirical genetic map. In each column of this figure
we show inferred $N(t)$ for the three extant populations in the model.
In each row we show comparisons among the methods (note that we compare two differing
sample sizes for \MSMC). The solid black lines represent the true effective population
size as inferred from the inverse coalescent rate defined by the model.
Blue lines represent estimates from each of three replicate whole genome simulations.
While there is variation in accuracy among methods, populations, and individual replicates,
generally the methods are accurate for this model of human history.

\stdpopsim allows us to readily compare these estimates to those drawn from a different
model of human history. In Figure \ref{fig:n_t_gutenkunst} we show estimates of
$N(t)$ from simulations using the same physical and genetic maps, but a different demographic
history---that inferred by \cite{gutenkunst2009inferring}. Again we see that each
of the methods is capturing relevant parts of the population history, although the
discrepancy between inferred and simulated values varies with $t$. In comparing inferences between the
models it is interesting to note that $N(t)$ estimates for the CHB and CEU
simulated populations are generally better across methods than estimates from the YRI
simulated population.

We can even do away with the demographic model entirely, and simulate under a human
genome architecture and genetic map to see how well methods might do at recovering
the population history of a constant sized population. We show results of such an
experiment in Figure \ref{fig:n_t_HomSap_constant}. Again we see that all methods
are recovering a modern population size within a factor of two of the truth, however
SMC-based methods, perhaps due to their regularization, tend to learn sinusoidal
patterns of population size even though no change is present.

@@

As most method development for population genetics has been focused on human-scale
data, it is of consequence to ask how such methods might perform in non-human
genomes. Figure \ref{fig:n_t_sheehan} shows parameter estimates from a demographic
model estimated from an African sample of \emph{Drosophila melanogaster} \citep{sheehan2016deep}.
As with humans, we use \stdpopsim to simulate replicate \emph{Drosophila} genomes under
an empirically derived genetic map, and then try to infer back the parameters of the model.
Accuracy is mixed among methods in this setting, and generally worse than what we
observe under simulations of the human genome.

\begin{figure}
\begin{center}
\includegraphics[width=0.7\linewidth]{display_items/homo_sapiens_two_popn_comp.png}
\caption{\textbf{Parameters estimated using a multi-population human model}.
Here we show estimates of $N(t)$ inferred using \dadi, \texttt{fastsimcoal2}, or \smcpp.
A) Data were generated by simulating
replicate human genomes under the \cite{gutenkunst2009inferring} model and using the genetic map
inferred in \cite{international2007second}. B) For \dadi and \texttt{fastsimcoal2} we show parameters inferred
by fitting the depicted IM model, which includes population sizes, migration rates, and a split
time between CEU and YRI samples. C) Population size estimates for each population (rows)
from \dadi, \texttt{fastsimcoal2}, and \smcpp (columns). In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model. The population split date, $T_{div}$, is shown at
the bottom, with a common X-axis to the background figure.}
\label{fig:IM_popn_human}
\end{center}
\end{figure}

\paragraph*{Multi-population demographic models}
As \texttt{stdpopsim} implements multi-population demographic models, we also
explored parameter estimation of population divergence parameters. In particular
we performed simulations under multi-population models for humans and \emph{Drosophila}
and inferred simulated parameters using \dadi, \fastsimcoal, and \smcpp.
For simplicity, we conducted inference in \dadi and \fastsimcoal by fitting an isolation with migration (IM) model
with constant population sizes and bi-directional migration. For human
models with more than two-populations, e.g. \cite{gutenkunst2009inferring},
this means that we are inferring parameters for a model which itself does
not match the model from which the data were generated (Figure
\ref{fig:IM_popn_human}A and B), however this is common practice in the field
when applying IM-type models.

In Figure \ref{fig:IM_popn_human}C we show estimates of population sizes and divergence
time, for each of the inference methods, using samples drawn from African and European populations
simulated under the \cite{gutenkunst2009inferring} model. Each of the methods compared has stronger
and weaker points to their accuracy across parameters. For instance the simple, IM-style
analyses we have performed do not capture recent population size change in CEU, however
they do a good job of estimating the simpler simulated YRI population size history from this model.

Again we can compare between species and look at the performance of these methods in
the context of a two-population model from \emph{Drosophila melanogaster}. Figure
\ref{fig:two_popn_fly} shows parameter estimates for simulations drawn from
the model of \cite{li2006inferring}, which includes
an ancestral population in Africa, then a population expansion with a population
split and bottleneck into a European population. These simulations are generated
using the recombination map of \cite{comeron2012many}. Here methods differ dramatically
in the estimates, with \smcpp yielding quite noisy estimates of population size
towards the more ancient past and all methods failing to capture the very brief
population bottleneck in Europe, although the IM model of course only allows for
two epochs in the population size history.

@@

The \stdpopsim library is the first piece of this effort, in that it allows for rigorous
standardization of complex population genetic simulations. Population genetics, as a field,
has yet to coalesce around a set of gold-standards for the crucial task of method
evaluation, which in our field hinges on simulation. In contrast, other fields such as
structural biology \citep{moult1995large} and machine learning \citep{russakovsky2015imagenet} have a long track record
of standardized method testing. We hope that our efforts represent the beginning of what
will prove to be an equally longstanding and invaluable tradition in population genetics.

@@

We have built \stdpopsim, and the PopSim Consortium, to be an open, community-developed
project, meaning that it has excellent potential for future growth.
Thus far we have implemented genome representations and genetic maps for the three most
common study systems in computational genetics: humans, \emph{Drosophila}, and \emph{Arabidopsis}.
The standard operating procedures that we have outlined in our detailed documentation
means that new researchers, who are interested in adding their favorite organism/genome
are free to do so with very few barriers to entry (see Methods).
Further, new demographic models can be added
to each of the organisms in the catalog through similarly well documented steps.
A long-term goal then is that through a mechanism such as \stdpopsim, the community
can determine (in an open way) what is the correct implementation for a published demographic model,
and in turn, what are the models in which the community is most interested.

@@

In the use case comparisons we have performed,
we found that the methods tested \emph{do} generally perform well on human genomes,
up to, for instance, the usual factor-of-two differences in inferred population sizes.
This will not be surprising to people familiar with the landscape of
demographic inference methods in population genetics,
but the nature and scale of uncertainty may be not clear to people new to the field.
The ability for \stdpopsim to generate data with and without significant features, such
as the genetic map or population size change (e.g., Figure \ref{fig:n_t_HomSap_constant}), allows
us to ask directly about the various failure modes of popular methods.
Moreover the comparison of methods across the various genome organizations, genetic maps,
and demographic histories of different organisms, provides valuable information
about how methods might perform on non-human systems.
In our initial comparison we find that inference of population histories from simulated \emph{Drosophila}
or \emph{Arabidopsis} genomes is a fair bit more difficult from SMC based methods
than in the human genome. Likely this is at least partly a consequence of the reduced physical and genetic
map lengths of these genomes, although we will leave further investigation of this to the future.
Finally, comparison of results across methods or simulation runs
gives a strong estimate of inference uncertainty, analogous to parametric
bootstrapping---although different methods have different strengths and weaknesses,
strong conclusions should probably not rest entirely on features inferred by only one method.

The broad brush comparisons we have done, while far from exhaustive, are informative,
and highlight the utility of easy comparison of methods on the same footing.
The tests here were under best-case scenarios---simulations of discrete,
randomly mating populations that fit the assumptions of the methods nearly
exactly---and yet different methods, as well as applications of the same method
to different simulations of the same model,
produced somewhat different results. Future efforts will extend the resource described here to
encompass additional population genetic tasks, such as detecting natural selection, inferring
recombination rates, etc, in hopes of producing a general-purpose resource for fair and
standardized evaluation of the ever-increasing arsenal of computational population genetic
inference tools.
